---
title: "Assignment - SQL and R"
author: "Vincent Miceli"
date: "2/9/2020"
output: pdf_document
---
# Overview

For this assignment, I created a table of movie ratings and stored them in a SQL database using PostgreSQL.  I created a table called ratings with columns Title, James, Lance, Lou, Maria, and Sarah, all of type character variable.  I then executed the following SQL statement to fill in the values:

INSERT INTO public.ratings ("Title", "James", "Lance", "Lou", "Maria", "Sarah")
VALUES 
	('The Lion King', 5, 4, 4, 3, ''),
	('Mulan', 4, 3, 5, '', 5),
	('Aladdin', '', 3, 4, 5, 2),
	('Pocahontas', 3, '', 4, 5, 4),
	('Hercules', 3, '', 4, 3, 3),
	('101 Dalmations', 3, 2, 5, '', 4);
	
I then exported the table to a csv file called ratings.csv.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

```{r}
ratings <- read.csv('ratings.csv')
```

```{r}
ratings
```

# Conclusion

The resulting data frame from importing the ratings.csv file is a simple table of movie ratings from 5 different people.  I included missing values in instances where the person had never seen the movie.

