---
title: "Assignment_1"
author: "Vincent Miceli"
date: "1/31/2020"
output: html_document
---
# Overview

Many sport leagues today are making use of data analytics to measure the value/impact a player has on the game.  Knowing this information allows teams to find the best players and try to acquire them, or give more opportunities to their own players.  A new metric for evaluating NBA players was recently made, called RAPTOR. The link to the article is below: 

https://fivethirtyeight.com/features/introducing-raptor-our-new-metric-for-the-modern-nba/

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

```{r}
raptor <- read.csv('latest_RAPTOR_by_player.csv')
```

```{r}
head(raptor)
```

```{r}
colnames(raptor)
```

```{r}
pacman::p_load(dplyr, tidyr, magrittr)
raptor %<>%
  select(-c(season, war_playoffs,
            war_reg_season))
```

```{r}
raptor %<>%
  rename(possessions = poss,
         minutes_played = mp)

colnames(raptor)
```
# Findings and Recommendations

The columns in RAPTOR are likely all significant in being able to measure a player's value, which is generalized by the column "war_total", so I did not remove many columns.  I removed season as all the data is from the same season, 2020.  I removed war_playoffs and war_reg_season because the playoffs have not happened yet and is just a column of 0's, all the information is based on regular season data.  I then renamed poss to possessions and mp to minutes played as they are not intuitive names.   

