---
title: "Assignment 7"
author: "Vincent Miceli"
date: "3/13/2020"
output: html_document
---

## Overview

I created the same table in three different file types, html, json, and xml.  The table contains the Title, Author, Year, and Genre of three particular books.  I then loaded the individual files into R as data frames.

```{r}
pacman::p_load(XML, jsonlite)
```


```{r}
books_html <- readHTMLTable('books.html')[[1]]
books_html
```

```{r}
books_json <- fromJSON('books.json')
books_json
```

```{r}
books_xml <- xmlToDataFrame('books.xml')
books_xml
```

# Files

<a href"file:///=C:\Users\micel\Documents\607\Assignment5\books.html">html file</a>
<a href"file:///=C:\Users\micel\Documents\607\Assignment5\books.json">json file</a>
<a href"file:///=C:\Users\micel\Documents\607\Assignment5\books.xml">xml file</a>

## Conclusion

All three of the data frames are identical.  Since I created them by hand with the same exact structure, it makes sense that they are identical.