---
title: "Assignment 3"
author: "Vincent Miceli"
date: "2/16/2020"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#1. Using the 173 majors listed in fivethirtyeight.com’s College Majors dataset [https://fivethirtyeight.com/features/the-economic-guide-to-picking-a-college-major/], provide code that identifies the majors that contain either "DATA" or "STATISTICS"

```{r}
data <- read.csv('majors-list.csv')
majors <- data$Major

DATA <- grep('DATA', majors, value = TRUE)
STATISTICS <- grep('STATISTICS', majors, value = TRUE)

data_statistics <- c(DATA, STATISTICS)
data_statistics
```


#2 Write code that transforms the data below:

[1] "bell pepper"  "bilberry"     "blackberry"   "blood orange"

[5] "blueberry"    "cantaloupe"   "chili pepper" "cloudberry"  

[9] "elderberry"   "lime"         "lychee"       "mulberry"    

[13] "olive"        "salal berry"

Into a format like this:

c("bell pepper", "bilberry", "blackberry", "blood orange", "blueberry", "cantaloupe", "chili pepper", "cloudberry", "elderberry", "lime", "lychee", "mulberry", "olive", "salal berry")

```{r}
pacman::p_load(stringr)

data <- '[1] "bell pepper"  "bilberry"     "blackberry"   "blood orange"

[5] "blueberry"    "cantaloupe"   "chili pepper" "cloudberry"  

[9] "elderberry"   "lime"         "lychee"       "mulberry"    

[13] "olive"        "salal berry"'

data <- str_extract_all(data, "(.+?)+")
data <- unlist(data)

data
```


The two exercises below are taken from R for Data Science, 14.3.5.1 in the on-line version:

#3 Describe, in words, what these expressions will match:

(.)\1\1

This matches a character followed by the characters "\1\1", such as "h\1\1".  It is not backreferenced.

"(.)(.)\\2\\1"

This matches substrings starting with any character followed by a character, followed by the second character, then the first character, such as anna or otto.

(..)\1

This matches any two characters followed by "\1", such as "hi\1".  It is not backreferenced.

"(.).\\1.\\1"

This matches substrings starting with any 2 characters followed by the first character, then any character, then the first character again, such as "papop".

"(.)(.)(.).*\\3\\2\\1"

This matches strings starting with 4 or more characters, followed by the 3rd, 2nd, and 1st character in that order, such as "moooooooooom".


#4 Construct regular expressions to match words that:

Start and end with the same character.

"(.).*\\1"

Contain a repeated pair of letters (e.g. "church" contains "ch" repeated twice.)

".*(..).*\\1.*"

Contain one letter repeated in at least three places (e.g. "eleven" contains three "e"s.)

".*(.).*\\1.*\\1.*"

