---
title: "Assignment 3"
author: "Vincent Miceli"
date: "2/16/2020"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#1. Using the 173 majors listed in fivethirtyeight.com’s College Majors dataset [https://fivethirtyeight.com/features/the-economic-guide-to-picking-a-college-major/], provide code that identifies the majors that contain either "DATA" or "STATISTICS"

```{r}
data <- read.csv('majors-list.csv')
majors <- data$Major

DATA <- grep('DATA', majors, value = TRUE)
STATISTICS <- grep('STATISTICS', majors, value = TRUE)

data_statistics <- c(DATA, STATISTICS)
data_statistics
```


#2 Write code that transforms the data below:

[1] "bell pepper"  "bilberry"     "blackberry"   "blood orange"

[5] "blueberry"    "cantaloupe"   "chili pepper" "cloudberry"  

[9] "elderberry"   "lime"         "lychee"       "mulberry"    

[13] "olive"        "salal berry"

Into a format like this:

c("bell pepper", "bilberry", "blackberry", "blood orange", "blueberry", "cantaloupe", "chili pepper", "cloudberry", "elderberry", "lime", "lychee", "mulberry", "olive", "salal berry")

```{r}
m1 <- matrix(data = c("bell pepper", "bilberry", "blackberry", "blood orange", "blueberry", "cantaloupe", "chili pepper", "cloudberry", "elderberry", "lime", "lychee", "mulberry", "olive", "salal berry"))

as.vector(m1)
```


The two exercises below are taken from R for Data Science, 14.3.5.1 in the on-line version:

#3 Describe, in words, what these expressions will match:

(.)\1\1

This matches a character followed by \1\1

"(.)(.)\\2\\1"

This matches a string starting with any character followed by a character, followed by the second character, then the first character

(..)\1

This matches any two characters followed by \1

"(.).\\1.\\1"

This matches strings starting with any 2 characters followed by the first character, then any character, then the first character again

"(.)(.)(.).*\\3\\2\\1"

This matches strings starting with 4 or more characters, followed by the 3rd, 2nd, and 1st character in that order


#4 Construct regular expressions to match words that:

Start and end with the same character.

(.).*\\1

Contain a repeated pair of letters (e.g. "church" contains "ch" repeated twice.)

.*(..).*\\1.*

Contain one letter repeated in at least three places (e.g. "eleven" contains three "e"s.)

.*(.).*\\1.*\\1.*

