---
title: "Assignment 10"
author: "Vincent Miceli"
date: "4/5/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

This code is taken from chapter 2 of Text Mining With R by Julia Silge and David Robinson.  I will apply the code to a different corpus of text and incorporate an additional lexicon.

```{r}
library(tidytext)
library(textdata)
library(janeaustenr)
library(dplyr)
library(tidyr)
library(stringr)

tidy_books <- austen_books() %>%
  group_by(book) %>%  
  mutate(linenumber = row_number(),chapter = cumsum(str_detect(text, regex("^chapter [\\divxlc]",ignore_case = TRUE)))) %>%
  ungroup() %>%
  unnest_tokens(word, text)
  
nrcjoy <- get_sentiments("nrc") %>%
  filter(sentiment == "joy")
  
janeaustensentiment <- tidy_books %>%  
  inner_join(get_sentiments("bing")) %>%  
  count(book, index = linenumber %/% 80, sentiment) %>%
  spread(sentiment, n, fill = 0) %>% 
  mutate(sentiment = positive - negative)

library(ggplot2)
ggplot(janeaustensentiment, aes(index, sentiment, fill = book)) +  
  geom_col(show.legend = FALSE) +  
  facet_wrap(~book, ncol = 2, scales = "free_x")
```
*    Title: Text Mining With R Chapter 2 code
*    Author: Silge, J
*    Date: 2017
*    Availability: http://dl.booktolearn.com/ebooks2/computer/programming/9781491981658_Text_Mining_with_R_f482.pdf

```{r}
jeopardy <- read.csv('JEOPARDY_CSV.csv')
head(jeopardy, 3)
```

This is a dataset of Jeopardy questions and answers from various shows, and includes information such as the type of round, the value of the question, the category, and more.

```{r}
question_tokens <- jeopardy %>%
  mutate(Question = as.character(Question)) %>%
  unnest_tokens(word, Question)
```

```{r}
question_sentiment <- question_tokens %>%  
  inner_join(get_sentiments("loughran")) %>%
  count(Answer, Round, sentiment) %>%
  spread(sentiment, n, fill = 0) %>% 
  mutate(sentiment = positive - negative) %>%
  arrange(Round) %>%
  mutate(index = c(1:sum(Round == 'Double Jeopardy!'), 1:sum(Round == 'Final Jeopardy!'), 1:sum(Round == 'Jeopardy!'), 1:sum(Round == 'Tiebreaker')))
```

This creates a similar data frame to the one we created from Jane Austen books, and we have it indexed by round so that we can plot it in a similar way, making a subplot for each round type. I used the loughran lexicon this time.

```{r}
ggplot(question_sentiment, aes(index, sentiment, fill = Round)) +  
  geom_col(show.legend = FALSE) +  
  facet_wrap(~Round, ncol = 2, scales = "free_x")
```

